Cryptanalysis of SBC based post-quantum signature scheme

### 2025-06-05 rapport.tex
rearrangement des sections

	\section{Preliminaries and notations}
		In this section, we introduce some standard notations of the field and notions that we will use.

@@

		\subsection{Gröbner basis theory}
		Gröbner basis theory was first introduced by Buchberger in his thesis \cite{Buc} and he gave an algorithm to compute the Gröbner basis of an ideal.
		Gröbner basis algorithms like Buchberger's Algorithm \cite{Buc}, are used to find solutions for polynomial systems using algebraic geometry tools. For a gentle introduction to this subject \cite{CLS}.

@@

	\section{SBC Signature Scheme}
		\subsection{SBC Problem}
		The problem is the following: let $q$ be a prime power and two positive integers $k$, $n$.
		\newline
		Given two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$, which are linearly independent over $\mathbb{F}_q$, find two non-colinear vectors $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ such that $$(\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) = (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x})$$
		where $(\cdot)$ is the dot product between 2 vectors.
		We denote an instance of the SBC problem given by two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$ by $\text{SBC}[\vec{u}, \vec{v}]$. If the vectors $\vec{x},\vec{y} \in (\mathbb{F}_q)^n$ are solution of $\text{SBC}[\vec{u},\vec{v}]$, we use the notation $(\vec{x}, \vec{y}) \in \text{SBC}[\vec{u}, \vec{v}]$.

		\section{Cryptanalysis of the scheme}
		One of the known attacks given by the authors is through algebraic cryptanalysis.
		One can model the attack by taking $g$ as the following polynomial
		$$
		g(x_1,...,x_n,y_1,...,y_n) := (\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) - (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x}).
		$$
		We consider
		$$
		g(x_1,..,x_{n-2},y_1,..,y_{n-2}) := (\sum_{i = 1}^{n-2}u_{i}x_{i} + u_{n-1})(\sum_{i = 1}^{n-2}v_{i}y_{i} + v_{n}) - (\sum_{i = 1}^{n-2}u_{i}y_{i} + u_{n})(\sum_{i = 1}^{n-2}v_{i}x_{i} + v_{n-1}).
		$$

@@

	\section{Boolean systems}
		Sometimes, the finite field used is $\mathbb{F}_2$ because it is still supposedly hard to solve but is very easy to use and manipulate on computers. But in such small fields one can begin to use exhaustive search to obtain the solution to the system. The best algorithm known to the authors for exhaustive search for multivariate polynomial system $f_1,\dots,f_m$ over $\mathbb{F}_2[x_1,\dots,x_n]$ is \cite{BCC+10} with a complexity $\mathcal{O}(ln(n)2^n)$. What is remarkable is that the complexity does not implicate the number of polynomials in the system, $m$. There is also a software made by one of the authors, Charles Bouillaguet, that is available online and easy to use, \textit{libFES-lite} \cite{Bouilla22} and there have been implementations on FPGA that broke real world cryptography \cite{DDVY20}..

@@

@@

	\section{Security analysis}
		This section is heavily inspired from the security analysis of the MQOM signature scheme \cite{MQOMv2}.

@@

		\subsection{Attacks against SBC signatures}
		First we need to enumerate the parameters for the signature scheme. The finite field used is $\mathbb{F}_2$ and the vector length $n$ is 130 but we consider the NSBC version so it is 128 in practice with the 2 last coordinates implicitly defined. The field extension used $k$ is 257 which is prime to avoid subfield attacks \cite{Peb24}.

@@

		\subsection{Memory footprint of the attacks}

>>>>

	\section{MPCitH and SBC signature scheme}
		In this section, we introduce some standard notations of the field and notions that we will use.

@@

		\subsection{SBC Problem}
		The problem is the following: let $q$ be a prime power and two positive integers $k$, $n$.
		\newline
		Given two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$, which are linearly independent over $\mathbb{F}_q$, find two non-colinear vectors $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ such that $$(\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) = (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x})$$
		where $(\cdot)$ is the dot product between 2 vectors.
		We denote an instance of the SBC problem given by two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$ by $\text{SBC}[\vec{u}, \vec{v}]$. If the vectors $\vec{x},\vec{y} \in (\mathbb{F}_q)^n$ are solution of $\text{SBC}[\vec{u},\vec{v}]$, we use the notation $(\vec{x}, \vec{y}) \in \text{SBC}[\vec{u}, \vec{v}]$.

	\section{Polynomial system solving}
		Gröbner basis theory was first introduced by Buchberger in his thesis \cite{Buc} and he gave an algorithm to compute the Gröbner basis of an ideal.
		Gröbner basis algorithms like Buchberger's Algorithm \cite{Buc}, are used to find solutions for polynomial systems using algebraic geometry tools. For a gentle introduction to this subject \cite{CLS}.

@@

@@

		\subsection{Boolean systems}
		Sometimes, the finite field used is $\mathbb{F}_2$ because it is still supposedly hard to solve but is very easy to use and manipulate on computers. But in such small fields one can begin to use exhaustive search to obtain the solution to the system. The best algorithm known to the authors for exhaustive search for multivariate polynomial system $f_1,\dots,f_m$ over $\mathbb{F}_2[x_1,\dots,x_n]$ is \cite{BCC+10} with a complexity $\mathcal{O}(ln(n)2^n)$. What is remarkable is that the complexity does not implicate the number of polynomials in the system, $m$. There is also a software made by one of the authors, Charles Bouillaguet, that is available online and easy to use, \textit{libFES-lite} \cite{Bouilla22} and there have been implementations on FPGA that broke real world cryptography \cite{DDVY20}..

@@

	\section{Contributions}

		\subsection{System Modeling}
		One of the known attacks given by the authors is through algebraic cryptanalysis.
		One can model the attack by taking $g$ as the following polynomial
		$$
		g(x_1,...,x_n,y_1,...,y_n) := (\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) - (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x}).
		$$
		We consider
		$$
		g(x_1,..,x_{n-2},y_1,..,y_{n-2}) := (\sum_{i = 1}^{n-2}u_{i}x_{i} + u_{n-1})(\sum_{i = 1}^{n-2}v_{i}y_{i} + v_{n}) - (\sum_{i = 1}^{n-2}u_{i}y_{i} + u_{n})(\sum_{i = 1}^{n-2}v_{i}x_{i} + v_{n-1}).
		$$

@@

@@

		\subsection{Security analysis}
		This section is heavily inspired from the security analysis of the MQOM signature scheme \cite{MQOMv2}.

		First we need to enumerate the parameters for the signature scheme. The finite field used is $\mathbb{F}_2$ and the vector length $n$ is 130 but we consider the NSBC version so it is 128 in practice with the 2 last coordinates implicitly defined. The field extension used $k$ is 257 which is prime to avoid subfield attacks \cite{Peb24}.

@@

		\subsection{Space complexity of Crossbred}
